Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of Complex Quasiperiodic Variables from TESS

### 2023-08-11 paper/ms.tex
ditto
ms: update

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MS TODO's
%
% * find/replace "centrifugally supported" with something less
% jargony, and more physically meaningful.  (read CR89)
%
% * "Bound-free absorption can provide
%   wavelength-dependent opacity, but the absorption coefficients
%   generally {\it grow} with increasing wavelength, rather than {\it
%   shrink} as one needs in order to get deeper absorption in the blue
%   than in the red."
%   --> Isn't this statement wavelength dependent????
%
% * Verify "32" in Figure 4 does not change with number of stars in
%   subsample changing
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We performed an independent search for strongly periodic phenomena
using our Fourier pipeline, following \citet{2019ApJ...876..127Z} and
\citet[][their Section~1.3]{2023MNRAS.524.4220P}.
Starting with the PDC\_SAP light curves, we normalized each light
curve, and then re-binned it into equal width bins {\bf LB: how long?} to
account for the uneven spacing in the TESS data, as well as the data
gap during each sector caused by satellite downlink.  We then padded
the data to ensure that the light curve had a length that was a power
of two, as described by \citeauthor{2019ApJ...876..127Z}.  We
then took the Fourier transform of the corrected, padded light curve using 
{\tt numpy.fft.fft} and searched for peaks with a significance of over 
12-$\sigma$ within each set of 500 bins.

@@

%TODO: include?
%Between the targets identified as CQVs by both pipelines,
% {\red 
%\begin{itemize}
%\item XX targets were found to overlap
%\item YY targets were found via Fourier but not via dip counting
%\end{itemize}
%}
%The difference in discovery rate between these techniques can be attributed to
%low-amplitude peaks in the periodogram that were missed, some of which likely
%arose from fainter CQVs.

@@

The mosaic in Figure~\ref{fig:cqvs} shows phased light curves for the
\ngoods\ CQVs.  The objects are sorted first in order of the
number of TESS 120-second cadence sectors in which they clearly
demonstrated the CQV phenomenon, and secondarily by descending
brightness.  The top five objects by this metric are TIC~300651846
($T$$=$13.5, 12 sectors); TIC~402980664 ($T$$=$11.1, 7 sectors);
TIC~89463560 ($T$$=$13.5, 5 sectors); TIC~363963079 ($T$$=$12.9, 5
sectors); and TIC~294328887 ($T$$=$14.2, 4 sectors).  The brightest
five CQVs span 9.3$<$$T$$<$11.1; the faintest five span
14.5$<$$T$$<$15.0.  The fastest five have periods spanning
3.6\,hr$<$$P$$<$6.2\,hr, and the slowest five span
27\,hr$<$$P$$<$38\,hr.
Regarding likely astrometric or visual binaries,
\ngoodhighruwe/\ngoods\ high-quality CQVs have a Gaia DR3 RUWE$>$2;
\nmaybehighruwe/\nmaybes\ of the ambiguous sample share this
characteristic.
% This is consistent with Rappaport2014's visual binary fraction determined from
% 30/100 rapidly rotating Kepler stars having multiple periodicities

@@

Of our \ncpvsfound\ confirmed and candidate CQVs, \nnotfieldbanyan\
were associated with a nearby moving group or open cluster using
BANYAN\,$\Sigma$.  The relevant groups are listed in
Table~\ref{tab:thetable}; their ages span $\approx$5-200\,Myr.  The
most prodigious groups were Sco-Cen, Tuc-Hor, and Columba.  Six CQVs
were also identified in the Argus association
\citep{2019ApJ...870...27Z}, which serves as an indirect line of
evidence supporting the reality, and youth, of that group.  The yield
in Sco-Cen is not particularly surprising, since Sco-Cen contains the
majority of pre-main-sequence stars in the solar neighborhood.
However, given the $\lesssim$$10\%$ completeness of TESS beyond
100\,pc, there are probably many more CQVs that remain to be
discovered in Sco-Cen.  
%IC\,2602 and IC\,2391, which are located slightly beyond
%100\,pc, are an example of this completeness effect: our search did
%not find any CQVs in those clusters.

@@

%FIXME it'd be useful to get an RV for this one...
The upper age limit for CQVs may even pass 200\,Myr, based on the
candidate membership of TIC~294328887 in the Carina Near moving group
\citep{2006ApJ...649L.115Z}.  This group's $200 \pm 50$\,Myr age
primarily rests on the lithium sequence of its G-dwarfs
\citep{2006ApJ...649L.115Z}, which shows a coeval population of stars
older than the Pleiades and younger than the 400\,Myr Ursa Major
moving group.  The formal BANYAN-$\Sigma$ membership probability is
however somewhat low, perhaps due to the missing radial velocity.
This could be relatively easily resolved by acquiring even a
medium-resolution spectrum.
An independent assessment of the group's kinematics using Gaia data,
and its rotation sequence using TESS, could lend further credence in
such an analysis.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f4.pdf}
		\vspace{-0.6cm}
		\caption{
      %TODO : 
		{\bf CQVs keep their periods but change their shapes.}
      Out of the \ngoods\ CQVs in Figure~\ref{fig:cqvs}, 32 had
      120-second cadence TESS data available for a baseline of at
      least two years; the 27 brightest are shown here due to space
      constraints.  Each panel shows one sector of TESS data, and is
      phased to its deepest minimum in flux.  Each panel's title shows
      the TIC identifier and approximate period in hours.  Text insets
      show the TESS sector numbers, which generally span two years, or
      at least 1{,}000 cycles.  The vertical scale is fixed across
      sectors to clarify shape changes.  Gray circles are raw 2-minute
      data; colored circles bin to 300 points per cycle. 
		}
		\label{fig:evoln}
	\end{center}
\end{figure*}

@@

The average period, determined by measuring the PDM peak period over
each sector independently, was $\langle P \rangle = 18.5560$\,hr.  The
range between the maximum and minimum sector-specific periods was
measured to be about one minute.   However, in detail, a period shift of
$\pm$1\,minute would yield major phase drifts over the baseline; that
time interval corresponds to roughly 1/1000$^{\rm th}$ of a period,
and we have observed 1500 cycles.  
%Assuming $\sigma_{\phi} \approx
%0.01$ and applying Equation~\ref{eq:periodprecision} yields an
%expected period precision $\sigma_{\rm P} \approx 0.45\,{\rm sec}
%\approx 1.2\times 10^{-4}\,{\rm hr}$. 
 By visually fine-tuning over a
grid in period, we eventually found $P=18.5611 \pm 0.0001$\,hr, which
seems to track certain features in the LP 12-502 light curve well over
the entire dataset.

@@

% 16 px vs 310 px for a full phase
%Based on our SED fitting $R_\star=0.369\,R_\odot$
%star, and from the isochrone mass, $M_\star=0.215\,M_\odot$ 
{\sc Dip durations}---The shortest dip duration for any of the
individual LP 12-502 dips seems to be $\approx$0.06\,$P$ $\approx$
1.08\,hr.  In comparison, using the stellar radius and mass derived in
Section~\ref{subsec:starprops}, the characteristic timescale $T_{\rm
dur} \equiv R_\star P / (\pi a)$ for the transit of a point-source at
corotation is 1.03\,hr ({\bf uncertainties?}).  This means that while
some of the LP 12-502 dips are sufficiently long to require structures
that are extended in orbital azimuth, the durations of other dips are
consistent with effective radii for the occulting material $R_{\rm
eff} \ll R_\star$.  This implies $a/R_\star \approx 5.8$ for this
material, and so the analogous timescale at the stellar surface is
about six times slower.

@@

\subsection{Typical and extreme CQVs}
\label{subsec:extreme}
Figure~\ref{fig:catalogscatter} shows some of the derived properties for our
CQV catalog, and compares them against both the target star sample (top
panels), and against stars in the Pleiades
\citep{2016AJ....152..114R}.  Typical CQV masses span
0.1-0.4\,$M_\odot$, and typical ages span 2-150\,Myr.  This
is a mass and age regime that includes fully convective stars
($M_\star$$\lesssim$0.25\,$M_\odot$), as well as stars with radiative cores
($M_\star$$\gtrsim$0.25\,$M_\odot$; \citealt{2018A&A...619A.177B}).  We found
no obvious morphology differences between the two classes.  In terms of their
rotation, relative to the Pleiades, the CQVs are among the more rapidly
rotating half of M dwarfs.  The dearth of $T$$>$15.5 CQVs in our sample,
despite our search extending to $T$$=$16, is a consequence of the 120-second
selection function, not detection sensitivity.

@@

\subsection{TIC~435903839}
%TODO: move these to their own ``debunked'' table, I think...
%TODO: along with the IR excess ones
TIC~435903839, which we originally classified as an ``ambiguous'' CQV, is an apparently
similar case.
We initially labeled this source as having a 10.82\,hr period, because this
period minimized the dispersion in the phase-folded light curve.
However closer inspection reveals that this source is a photometric
blend of two ordinary rotating stars with $P_0$=3.60\,hr, and $P_1$=5.41\,hr.
This is a novel false positive scenario for CQVs:
two rapidly-rotating stars with near-integer ratios of rotation periods.
The beat between the two rotation signals produces the spurious
CQV signal.
Such false positives can be excluded through careful accounting of
all peaks in a periodogram.
For instance, TIC~435903839 shows a peak at 16.27\,hr, which is not
an integer multiple of the nominal ``10.82\,hr'' period.

\subsection{CQVs are quasiperiodic}
A periodic signal repeats exactly; the CQVs do not
(Figure~\ref{fig:evoln}).  While their periods appear to remain
constant to within measurement precision over thousands of cycles, the
light curve shapes evolve over 10 to 1{,}000 cycles.  They are
therefore {\it quasi}periodic.  This observation is consistent studies
by \citet{2022AJ....163..144G} and \citet{2023ApJ...945..114P}.
However it marks a qualitative departure from the ``persistent'' {\it
vs.} ``transient'' flux dip distinction described by
\citet{2017AJ....153..152S};  since all CQV dips seem to be transient
over timescales of more than 1{,}000 cycles (Figure~\ref{fig:evoln}).

@@

\subsection{Dip asymmetries and dust geometries}
The asymmetry of a dip can help diagnose the optical depth of the
occulting material as a function of orbital phase angle.  Sharp
leading edges with trailing exponential egresses for instance have
been previously seen for transiting exocomets and disintegrating rocky
bodies
\citep[e.g.][]{2012ApJ...752....1R,2012A&A...545L...5B,2015Natur.526..546V,2019A&A...625L..13Z}.

@@

\subsection{Dust, or gas?}
We believe that the most likely scenarios to explain CQVs are either
the dust clump scenario or the prominence scenario
(Figure~\ref{fig:f1}).  Both invoke clumpy material that would be
centrifugally supported at the corotation radius.  The prominence idea
has a longer history, based on analogy with quiescent
prominences/filaments observed to exist in the solar corona for up to
a few weeks \citep[see][]{2015ASSL..415.....V}.  In an extrasolar
context, spectroscopic detections of transient Balmer- and
resonance-line absorption seen for stars such as AB~Dor and Speedy~Mic
\citep[e.g.][]{1989MNRAS.238..657C,1993MNRAS.262..369J,2006MNRAS.365..530D,2016MNRAS.463..965L}
led to the interpretation that the data could best be explained by
similar structures: cold, minimally ionized hydrogen clouds or
filaments that scatter chromospheric emission from the star to produce
the observed spectroscopic line variations
\citep[see][]{1989MNRAS.238..657C}.  The short-term mechanical
stability of such gas configurations is theoretically plausible
\citep{2000MNRAS.316..647F,2022MNRAS.514.5465W}, and the
interpretation of this class of observations seems at least somewhat
secure.

@@

%TODO: improve this to make it less jargony
The difference between the prominence and dust clump scenarios is
essentially only in whether the occulting material of interest is
neutral hydrogen, or dust.  In the phrasing of the ``frozen flux''
condition of ideal rigid field magnetohydrodynamics, the tendency of
both to become trapped at the corotation radius in the equatorial
plane is tied to how of the four relevant forces (gravity, Lorentz,
inertial Coriolis, and inertial centrifugal), the Lorentz and Coriolis
only act perpendicular to field lines, while gravity and the
centrifugal force are in balance at $R_{\rm cr}$
\citep[see][Sec.~2]{2005MNRAS.357..251T}.  The magnetic field strength
is only relevant in this formulation of the system in that we must
have $R_{\rm sonic} < R_{\rm cr} < R_{\rm Alfven}$ in order for closed
loops to exist that can support prominences
\citep{2019MNRAS.482.2853J} {\bf lgb todo: verify \& de-jargonify!}.
In detail however, whether such magnetic fixed points lie in the
equatorial plane, or elsewhere, depends on the star's magnetic field
geometry \citep{2023MNRAS.518.4734S}.

\subsubsection{Gas absorption microphysics}
CQVs show broadband flux variations that can be 1-2$\times$ deeper in
the blue than in the red
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
For neutral hydrogen to produce this effect in absorption, its opacity
must be larger in the blue than in the red.
\citet{1992oasp.book.....G} suggests that this might be
microphysically possible: while bound-bound absorption provides
opacity only at narrow resonant lines, the hydrogen opacity due to
bound-free absorption is ``jagged'' \citep[see][Figure 8.5 and
Eq.~8.8]{1992oasp.book.....G}, such that at temperatures of
$\approx$3{,}000\,K to $\approx$10{,}000\,K the observed chromatic
transit depths might be reproduced.  Thompson scattering is ruled out
as a relevant opacity source, because it is gray.  The other plausible
alternative opacity source is dust, which also has a larger absorption
cross-section in the blue than the red \citep{1989ApJ...345..245C}.

@@

%\subsubsection{Why corotation?}
%All of the prominence models, and the diagrams of where the magnetic
%fixed points exist, enable the fixed points to not be located at
%corotation.
%In addition, the Balmer absorption line measurements have been used to
%argue for material being generally located at positions other than
%just corotation (CHECK Collier-Cameron1989).

\subsubsection{The lifetime constraint}
Independent of the microphysical basis for gas absorption, the
observed lifetime of the CQV phenomenon could provide another
dimension to discern between the gas {\it vs.} dust clump scenarios.
Based on the available statistics from e.g.
\citet{2022AJ....164...80R} and references therein, it seems plausible
that CQV occurrence decreases in time, bottoming to zero well before
the $\approx$700\,Myr age of Praesepe.  This is odd in the context of
the prominence scenario, because
pre-main-sequence M dwarfs spin up over the first $\approx$$10^8$\,yr;
prominences might even be expected to be {\it more} common at the age
of the Pleiades than for younger stars.  This broadly assumes that any
star that can support prominences will, and that the magnetic topology
of rapidly rotating M dwarfs at 100\,Myr is similar to that of 
rapidly rotating M dwarfs at say, 10\,Myr.  Based on the degree to
which M dwarfs remain rapid rotators over the first few gigayears
of their lives
\citep[e.g.][]{2022AJ....164...80R,2022ApJ...936..109P}, it seems surprising
in the gas clump (prominence) scenario that very few stars older than
150\,Myr have been seen to exhibit the phenomenon.  In the dust clump
scenario this is not a problem, because if the dust were externally
sourced, it would be expected to have a finite supply.

%\subsubsection{Does the CQV lifetime inform the mechanism?}
%Independent of the microphysical basis for gas absorption, the
%prominence scenario may present another issue: why don't stars in the
%Hyades and Praesepe show the phenomenon?
%In a broad sense, the prominence scenario argues that an intrinsic
%stellar process is key for generating the CQV phenomenon.
%If this were true, then stars of identical mass, size, and rotation
%periods might be expected to show the phenomenon in equal number,
%simply because the stellar dynamo is generated by fluid
%motions inside the star, which should be identical for stars of the
%same mass, size, and rotation period.
%
%Between ages of $\approx$100\,Myr (AB Dor, Pleiades, Psc-Eri) and
%$\approx$700\,Myr (Praesepe and Hyades), a 0.3\,M$_\odot$ star will
%shrink by $\approx$10\%, from $\approx$0.33\,R$_\odot$ to 
%$\approx$0.29\,R$_\odot$ \citep[e.g.][]{2014AJ....147...82C}.
%This is not a particularly drastic size change.
%Similarly, at 700\,Myr many M dwarfs have comparably rapid rotation periods 
%to those at 100\,Myr \citep[e.g. Figure~13 of][]{2022AJ....164...80R};
%the mean population does show evidence for spin-down after the
%PMS contraction finishes, but there are still a large number of 
%700\,Myr M dwarfs with $P_{\rm rot}\lesssim$2\,days, beyond
%which the CQV phenomenon becomes less common.
%
%Broadly speaking -- the absence of old CQVs seems to be
%more plausibly linked to a gradual depletion of dusty detritus from the
%planet formation era than it does to the changing stellar properties
%as the stars gradually finish their pre-main-sequence evolution, and then
%continue to spin rapidly for many hundreds of millions of years.

@@

Close-in planets are common around M-dwarfs;
studies from Kepler have shown that high-mass
M  dwarfs have $\approx$0.1 planets per star with sizes between
1-4\,$R_\oplus$ and orbital periods within 3 days
\citep{2015ApJ...807...45D}.  The number increases to $\approx$0.7
planets per star considering planets out to $P$$<$10\,days.
Extrapolating to all small close-in planets with say
0.1-4\,$R_\oplus$ and within 10 days, it is reasonable to expect on
average one close-in planet per M dwarf.  
%TRAPPIST-1b ($P$=1.5\,days) is an
%example of this type of planet, orbiting around a 0.08\,$M_\odot$ star
%\citep{2017Natur.542..456G}.

@@

In systems like LP 12-502, the observed dip depth variations would
necessarily drive one toward a picture of a disintegrating
planetesimal swarm, rather than just a single planetesimal.  
%While the number of free parameters in this type of model is somewhat
%dizzying, we have no right to believe that nature need be simple.
The need for a swarm rather than just one launching body would
be driven by the observation that we see dips appear and disappear at
many {\it different phases}, but with roughly the same period.
The observed changing dip depths in response to flares would be
understood through the planetesimal outflow rates being stalled by
changes in the local magnetic field or coronal plasma density.
The sizes of these purported planetesimals would need to be $\lesssim
1$\,$R_\oplus$, based on the non-detections of their transits,
analogous to e.g.  K2-22 \citep{2015ApJ...812..112S} or KOI-2700
\citep{2014ApJ...784...40R}.  Typical sizes of bodies in the asteroid
belt for instance span hundreds of meters to a few kilometers.

@@

%The dip-profile asymmetries during certain segments of the light curve
%plausibly match this idea; for instance, the complex in cycles 248-298
%could be well-fit by a sharp leading edge that decays over 0.2\,$P$
%(3.7\,hours).  The earlier complex, during cycles 248-264
%(pre-state-switch), could be fit by a similar profile.

%There are two degenerate explanations for the small
%``wiggles'' in the exponential decay.  In one explanation, the decay
%timescale is fast ($\approx$0.03\,$P$), and there are up to five
%exocomets, superposed.  In the other, there is only one main
%``launching body'', and the material it ejects either has some
%sub-structure, or else is passing across (and in synchronicity with) a
%star with heavily substructured starspots across its surface.

%Every dip in such a model would have as free parameters the dip phase,
%the duration (captured by an exponential decay timescale), and the dip
%depth.  In detail, the impact parameter would also enter.  Allotting
%four free parameters per dip yields up to 32 free parameters for the
%light curve at the most complicated times (assigning one exocomet per
%local minimum).

%It seems
%challenging to reconcile this behavior with the possible presence of
%just a single launching body.  Finally, the fact that the dips often
%respond to events like flares suggests that that the responsible
%material be much less massive than a 1--100\,km sized planetesimal,
%and therefore more easily influenced by the stellar magnetic field.

%  % TODO: needed??
%  {\sc Some CQVs require external occulters}
%  (rehash Bodman2017 timescale arguments)

%{\sc Each CQV a snowflake?}
%An intriguing aspect of Figure~\ref{fig:cqvs} is that very few CQVs
%exhibit exactly identical light curve morphology.
%There are of course similarities.
%TIC 272248916 and TIC 442571495 match.
%But overall, they all different.

% \paragraph{Predicting the future, given the past}
% Figure~\ref{fig:evoln}
% shows that for 21 CQVs that were observed roughly two years
% apart, 5 showed obvious signs of the phenomenon in one sector,
% and marginal or non-existence signs in the other.
% We emphasize that the periods themselves were usually stable to $<$0.1\%.
% However, this shape evolution
% suggests that CQVs have a duty cycle of $\approx$75\%.

% \paragraph{Why would the prominence scenario tend toward corotation?}
% This is analogous to quiescent
% prominences on the Sun \citep{1967SoPh....2...39K}.

>>>>

@@

We performed an independent search for strongly periodic phenomena
using a Fourier-based approach, following \citet{2019ApJ...876..127Z}
and \citet[][their Section~1.3]{2023MNRAS.524.4220P}.  Starting with
the PDC\_SAP light curves, we normalized each light curve, and then
re-binned it into equal width bins {\bf LB: how long?} to account for
the uneven spacing in the TESS data, as well as the data gap during
each sector caused by satellite downlink.  We then padded the data to
ensure that the light curve had a length that was a power of two, as
described by \citeauthor{2019ApJ...876..127Z}.  We then took the
Fourier transform of the corrected, padded light curve using {\tt
numpy.fft.fft} and searched for peaks with a significance of over
12-$\sigma$ within each set of 500 bins.

@@

@@

The mosaic in Figure~\ref{fig:cqvs} shows phased light curves for the
\ngoods\ CQVs.  The objects are sorted first in order of the number of
TESS 120-second cadence sectors in which they clearly demonstrated the
CQV phenomenon, and secondarily by descending brightness.  The top
five objects by this metric are TIC~300651846 ($T$$=$13.5, 12
sectors); TIC~402980664 ($T$$=$11.1, 7 sectors); TIC~89463560
($T$$=$13.5, 5 sectors); TIC~363963079 ($T$$=$12.9, 5 sectors); and
TIC~294328887 ($T$$=$14.2, 4 sectors).  The brightest five CQVs span
9.3$<$$T$$<$11.1; the faintest five span 14.5$<$$T$$<$15.0.  The
fastest five have periods spanning 3.6\,hr$<$$P$$<$6.2\,hr, and the
slowest five span 27\,hr$<$$P$$<$38\,hr.  Regarding likely astrometric
or visual binaries, \ngoodhighruwe/\ngoods\ high-quality CQVs have a
Gaia DR3 RUWE$>$2; \nmaybehighruwe/\nmaybes\ of the ambiguous sample
share this characteristic.

% This is consistent with Rappaport2014's visual binary fraction
% determined from 30/100 rapidly rotating Kepler stars having multiple
% periodicities

@@

Of our \ncpvsfound\ confirmed and candidate CQVs, \nnotfieldbanyan\
were associated with a nearby moving group or open cluster using
BANYAN\,$\Sigma$.  The relevant groups are listed in
Table~\ref{tab:thetable}; their ages span $\approx$5-200\,Myr.  The
most prodigious groups were Sco-Cen, Tuc-Hor, and Columba.  Six CQVs
were also identified in the Argus association
\citep{2019ApJ...870...27Z}, which serves as an indirect line of
evidence supporting the reality, and youth, of that group.  The yield
in Sco-Cen is not particularly surprising, since Sco-Cen contains the
majority of pre-main-sequence stars in the solar neighborhood.
However, given the $\lesssim$$10\%$ completeness of TESS beyond
100\,pc, there are probably many more CQVs that remain to be
discovered in Sco-Cen.

% IC\,2602 and IC\,2391, which are located slightly beyond 100\,pc, are
% an example of this completeness effect: our search did not find any
% CQVs in those clusters.

@@

% NOTE: it'd be useful to get an RV for this one...

The upper age limit for CQVs may even pass 200\,Myr, based on the
candidate membership of TIC~294328887 in the Carina Near moving group
\citep{2006ApJ...649L.115Z}.  This group's $200 \pm 50$\,Myr age
primarily rests on the lithium sequence of its G-dwarfs
\citep{2006ApJ...649L.115Z}, which shows a coeval population of stars
older than the Pleiades and younger than the 400\,Myr Ursa Major
moving group.  The formal BANYAN-$\Sigma$ membership probability is
however somewhat low, perhaps due to the missing radial velocity.
This could be relatively easily resolved by acquiring even a
medium-resolution spectrum.  An independent assessment of the group's
kinematics using Gaia data, and its rotation sequence using TESS,
could lend further credence in such an analysis.

@@

\begin{figure*}[!tp]
	\begin{center}
		\centering
		\includegraphics[width=\textwidth]{f4.pdf}
		\vspace{-0.6cm}
		\caption{
		{\bf CQVs keep their periods but change their shapes.}
      Out of the \ngoods\ CQVs in Figure~\ref{fig:cqvs}, 32 had
      120-second cadence TESS data available for a baseline of at
      least two years; the 27 brightest are shown here due to space
      constraints.  Each panel shows one sector of TESS data, and is
      phased to its deepest minimum in flux.  Each panel's title shows
      the TIC identifier and approximate period in hours.  Text insets
      show the TESS sector numbers, which generally span two years, or
      at least 1{,}000 cycles.  The vertical scale is fixed across
      sectors to clarify shape changes.  Gray circles are raw 2-minute
      data; colored circles bin to 300 points per cycle. 
		}
		\label{fig:evoln}
	\end{center}
\end{figure*}

@@

The average period, determined by measuring the PDM peak period over
each sector independently, was $\langle P \rangle = 18.5560$\,hr.  The
range between the maximum and minimum sector-specific periods was
measured to be about one minute.   However, in detail, a period shift
of $\pm$1\,minute would yield major phase drifts over the baseline;
that time interval corresponds to roughly 1/1000$^{\rm th}$ of a
period, and we have observed 1500 cycles.  By visually fine-tuning
over a grid in period, we eventually found $P=18.5611 \pm 0.0001$\,hr,
which seems to track certain features in the LP 12-502 light curve
well over the entire dataset.

@@

% 16 px vs 310 px for a full phase
% Based on our SED fitting $R_\star=0.369\,R_\odot$ star, and from the
% isochrone mass, $M_\star=0.215\,M_\odot$

{\sc Dip durations}---The shortest dip duration for any of the
individual LP 12-502 dips seems to be $\approx$0.06\,$P$ $\approx$
1.08\,hr.  In comparison, using the stellar radius and mass derived in
Section~\ref{subsec:starprops}, the characteristic timescale $T_{\rm
dur} \equiv R_\star P / (\pi a)$ for the transit of a point-source at
corotation is 1.03\,hr ({\bf uncertainties?}).  This means that while
some of the LP 12-502 dips are sufficiently long to require structures
that are extended in orbital azimuth, the durations of other dips are
consistent with effective radii for the occulting material $R_{\rm
eff} \ll R_\star$.  This implies $a/R_\star \approx 5.8$ for this
material, and so the analogous timescale at the stellar surface is
about six times slower.

@@

\subsection{Typical and extreme CQVs}
\label{subsec:extreme}

Figure~\ref{fig:catalogscatter} shows some of the derived properties
for our CQV catalog, and compares them against both the target star
sample (top panels), and against stars in the Pleiades
\citep{2016AJ....152..114R}.  Typical CQV masses span
0.1-0.4\,$M_\odot$, and typical ages span 2-150\,Myr.  This is a mass
and age regime that includes fully convective stars
($M_\star$$\lesssim$0.25\,$M_\odot$), as well as stars with radiative
cores ($M_\star$$\gtrsim$0.25\,$M_\odot$;
\citealt{2018A&A...619A.177B}).  We found no obvious morphology
differences between the two classes.  In terms of their rotation,
relative to the Pleiades, the CQVs are among the more rapidly rotating
half of M dwarfs.  The dearth of $T$$>$15.5 CQVs in our sample,
despite our search extending to $T$$=$16, is a consequence of the
120-second selection function, not detection sensitivity.

@@

\subsection{TIC~435903839}

%TODO: move these to their own ``debunked'' table, I think...
%TODO: along with the IR excess ones

TIC~435903839, which we originally classified as an ``ambiguous'' CQV,
is an apparently similar case.  We initially labeled this source as
having a 10.82\,hr period, because this period minimized the
dispersion in the phase-folded light curve.  However closer inspection
reveals that this source is a photometric blend of two ordinary
rotating stars with $P_0$=3.60\,hr, and $P_1$=5.41\,hr.  This is a
novel false positive scenario for CQVs: two rapidly-rotating stars
with near-integer ratios of rotation periods.  The beat between the
two rotation signals produces the spurious CQV signal.  Such false
positives can be excluded through careful accounting of all peaks in a
periodogram.  For instance, TIC~435903839 shows a peak at 16.27\,hr,
which is not an integer multiple of the nominal ``10.82\,hr'' period.

\subsection{CQVs are quasiperiodic}

A periodic signal repeats exactly; the CQVs do not
(Figure~\ref{fig:evoln}).  While their periods appear to remain
constant to within measurement precision over thousands of cycles, the
light curve shapes evolve over 10 to 1{,}000 cycles.  They are
therefore {\it quasi}periodic.  This observation is consistent studies
by \citet{2022AJ....163..144G} and \citet{2023ApJ...945..114P}.
However it marks a qualitative departure from the ``persistent'' {\it
vs.} ``transient'' flux dip distinction described by
\citet{2017AJ....153..152S};  since all CQV dips seem to be transient
over timescales of more than 1{,}000 cycles (Figure~\ref{fig:evoln}).

@@

\subsection{Dip asymmetries and dust geometries}

The asymmetry of a dip can help diagnose the optical depth of the
occulting material as a function of orbital phase angle.  Sharp
leading edges with trailing exponential egresses for instance have
been previously seen for transiting exocomets and disintegrating rocky
bodies
\citep[e.g.][]{2012ApJ...752....1R,2012A&A...545L...5B,2015Natur.526..546V,2019A&A...625L..13Z}.

@@

\subsection{Dust, or gas?}

We believe that the most likely scenarios to explain CQVs are either
the dust clump scenario or the prominence scenario
(Figure~\ref{fig:f1}).  Both invoke clumpy material that would be
centrifugally supported at the corotation radius.  The prominence idea
has a longer history, based on analogy with quiescent
prominences/filaments observed to exist in the solar corona for up to
a few weeks \citep[see][]{2015ASSL..415.....V}.  In an extrasolar
context, spectroscopic detections of transient Balmer- and
resonance-line absorption seen for stars such as AB~Dor and Speedy~Mic
\citep[e.g.][]{1989MNRAS.238..657C,1993MNRAS.262..369J,2006MNRAS.365..530D,2016MNRAS.463..965L}
led to the interpretation that the data could best be explained by
similar structures: cold, minimally ionized hydrogen clouds or
filaments that scatter chromospheric emission from the star to produce
the observed spectroscopic line variations
\citep[see][]{1989MNRAS.238..657C}.  The short-term mechanical
stability of such gas configurations is theoretically plausible
\citep{2000MNRAS.316..647F,2022MNRAS.514.5465W}, and the
interpretation of this class of observations seems at least somewhat
secure.

@@

%TODO: improve this to make it less jargony

The difference between the prominence and dust clump scenarios is
essentially only in whether the occulting material of interest is
neutral hydrogen, or dust.  In the phrasing of the ``frozen flux''
condition of ideal rigid field magnetohydrodynamics, the tendency of
both to become trapped at the corotation radius in the equatorial
plane is tied to how of the four relevant forces (gravity, Lorentz,
inertial Coriolis, and inertial centrifugal), the Lorentz and Coriolis
only act perpendicular to field lines, while gravity and the
centrifugal force are in balance at $R_{\rm cr}$
\citep[see][Sec.~2]{2005MNRAS.357..251T}.  The magnetic field strength
is only relevant in this formulation of the system in that we must
have $R_{\rm sonic} < R_{\rm cr} < R_{\rm Alfven}$ in order for closed
loops to exist that can support prominences
\citep{2019MNRAS.482.2853J} {\bf lgb todo: verify \& de-jargonify!}.
In detail however, whether such magnetic fixed points lie in the
equatorial plane, or elsewhere, depends on the star's magnetic field
geometry \citep{2023MNRAS.518.4734S}.

\subsubsection{Gas absorption microphysics}

CQVs show broadband flux variations that can be 1-2$\times$ deeper in
the blue than in the red
\citep{2020AJ....160...86B,2022AJ....163..144G,2023MNRAS.518.2921K}.
For neutral hydrogen to produce this effect in absorption, its opacity
must be larger in the blue than in the red.
\citet{1992oasp.book.....G} suggests that this might be
microphysically possible: while bound-bound absorption provides
opacity only at narrow resonant lines, the hydrogen opacity due to
bound-free absorption is ``jagged'' \citep[see][Figure 8.5 and
Eq.~8.8]{1992oasp.book.....G}, such that at temperatures of
$\approx$3{,}000\,K to $\approx$10{,}000\,K the observed chromatic
transit depths might be reproduced.  Thompson scattering is ruled out
as a relevant opacity source, because it is gray.  The other plausible
alternative opacity source is dust, which also has a larger absorption
cross-section in the blue than the red \citep{1989ApJ...345..245C}.

@@

%\subsubsection{Why corotation?}
%All of the prominence models, and the diagrams of where the magnetic
%fixed points exist, enable the fixed points to not be located at
%corotation.  In addition, the Balmer absorption line measurements
%have been used to argue for material being generally located at
%positions other than just corotation (CHECK Collier-Cameron1989).

\subsubsection{The lifetime constraint}

Independent of the microphysical basis for gas absorption, the
observed lifetime of the CQV phenomenon could provide another
dimension to discern between the gas {\it vs.} dust clump scenarios.
Based on the available statistics from e.g.
\citet{2022AJ....164...80R} and references therein, it seems plausible
that CQV occurrence decreases in time, bottoming to zero well before
the $\approx$700\,Myr age of Praesepe.  This is odd in the context of
the prominence scenario, because pre-main-sequence M dwarfs spin up
over the first $\approx$$10^8$\,yr; prominences might even be expected
to be {\it more} common at the age of the Pleiades than for younger
stars.  This broadly assumes that any star that can support
prominences will, and that the magnetic topology of rapidly rotating M
dwarfs at 100\,Myr is similar to that of rapidly rotating M dwarfs at
say, 10\,Myr.  Based on the degree to which M dwarfs remain rapid
rotators over the first few gigayears of their lives
\citep[e.g.][]{2022AJ....164...80R,2022ApJ...936..109P}, it seems
surprising in the gas clump (prominence) scenario that very few stars
older than 150\,Myr have been seen to exhibit the phenomenon.  In the
dust clump scenario this is not a problem, because if the dust were
externally sourced, it would be expected to have a finite supply.

@@

Close-in planets are common around M-dwarfs; studies from Kepler have
shown that high-mass M  dwarfs have $\approx$0.1 planets per star with
sizes between 1-4\,$R_\oplus$ and orbital periods within 3 days
\citep{2015ApJ...807...45D}.  The number increases to $\approx$0.7
planets per star considering planets out to $P$$<$10\,days.
Extrapolating to all small close-in planets with say 0.1-4\,$R_\oplus$
and within 10 days, it is reasonable to expect on average one close-in
planet per M dwarf.

%TRAPPIST-1b ($P$=1.5\,days) is an
%example of this type of planet, orbiting around a 0.08\,$M_\odot$ star
%\citep{2017Natur.542..456G}.

@@

In systems like LP 12-502, the observed dip depth variations would
necessarily drive one toward a picture of a disintegrating
planetesimal swarm, rather than just a single planetesimal.  The need
for a swarm rather than just one launching body would be driven by the
observation that we see dips appear and disappear at many {\it
different phases}, but with roughly the same period.  The observed
changing dip depths in response to flares would be understood through
the planetesimal outflow rates being stalled by changes in the local
magnetic field or coronal plasma density.  The sizes of these
purported planetesimals would need to be $\lesssim 1$\,$R_\oplus$,
based on the non-detections of their transits, analogous to e.g.
K2-22 \citep{2015ApJ...812..112S} or KOI-2700
\citep{2014ApJ...784...40R}.  Typical sizes of bodies in the asteroid
belt for instance span hundreds of meters to a few kilometers.

@@
